---
title: "Sound stats"
author: "Martin R Vasilev"
date: "27 September 2017"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = getwd())
```


## Data pre-processing:

```{r, echo= FALSE}

office= T

#source("https://raw.githubusercontent.com/martin-vasilev/R_scripts/master/useful_functions.R")

if(office){
    load("C:/Users/mvasilev/Documents/DEVS/preproc/sound_check.Rda")
} else{
    load("C:/Users/Martin Vasilev/Documents/DEVS/preproc/sound_check.Rda")
}


nobs<- nrow(sound_check)

# remove blinks on target words:
blinks<- which(sound_check$blink=='Yes')
nblinks<- length(blinks)
sound_check<- sound_check[-blinks,]


# remove sounds played after fixation has started:
infix<- which(sound_check$delFix>14)
infixn<- length(infix)
sound_check<- sound_check[-infix,]

#cat(sprintf("%d percent of data excluded due to blinks", (nblinks/nobs)*100))
#cat(sprintf("%d percent of data excluded due to in-fixations", (infixn/nobs)*100))
#cat(sprintf("%d percent of data remains for analysis", (nrow(sound_check)/nobs)*100))

sound_check<- subset(sound_check, delFix<80)

#sound_check<- subset(sound_check, sub!=5)

```
- based on `r length(unique(sound_check$sub))` subjects
- `r (nblinks/nobs)*100` % of data was excluded due to blinks on target words
- `r (infixn/nobs)*100` % of data was excluded due to late triggering of the sound (command sent after start of fixation)
- `r (nrow(sound_check)/nobs)*100` % of data remains for analysis 
- Number of observations per sound condition (overall): Deviant (`r table(sound_check$sound_type)[1]`), Standard (`r table(sound_check$sound_type)[3]`), Silence (`r table(sound_check$sound_type)[2]`)


## Timing of sound:

Note that technology doesn't allow absolute time-locking to the onset of fixation. The reason for this is that Eyelink's algorithm for online parsing of fixations requires about 35 ms to obtain enough samples to detect the start of a fixation. Therefore, if you wait for this flag from the system, the fixation would have already started ~ 35 ms ago. This is clearly too much delay for the present study. The best approach (adopted here) is to trigger the sound as soon as the eye crosses the start of the empty space before the word. Usually, this happens towards the end of the saccade.

The sound in this experiment requires 14 ms from sending the 'play' command to the signal coming out from the speakers. Since the command to play the sound is sent several ms before the start of fixation on the target word, this means that participants will usually hear the sound within several ms after the start of fixation. In other words, the latency between the start of fixation on the target word and participants hearing the sound is: 14 ms - time period between sending the command and the start of fixation. The distribution of this latency is shown below:


```{r, echo= FALSE}
hist(sound_check$delFix, xlab= "Sound onset delay (in ms) relative to fixation onset of the target word", main= "Sound implementation timing", family= "serif", col="lightgreen", freq=FALSE, breaks=30, xlim= c(-40, 20)) # , xlim= c(-20, 20)
m<- subset(sound_check, delFix>-40)
abline(v = mean(m$delFix, na.rm=T), col= "darkred", lwd=3)
curve(dnorm(x, mean=mean(sound_check$delFix, na.rm=T), sd=sd(sound_check$delFix, na.rm=T)), add=TRUE, col="darkblue", lwd=2) 

s2<- subset(sound_check, sound_type!="SLC")
```
*Note*: Extreme values due to long saccades are removed from this graph (`r round((length(which(sound_check$delFix< -40))/ nrow(sound_check))*100, 1)` %).

## Inter-stimulus Interval (ISI):

The pace of the sound presentation is determined by participants, in the sense that it depends on how long they fixate words in the sentence. For this reason, it's useful to know what is the average interval between playing two consecutive sounds: Mean: `r round(mean(s2$ISI-50, na.rm=T))`. SD: `r round(sd(s2$ISI-50, na.rm=T))`. Range: `r range(s2$ISI-50, na.rm=T)`. 

*Note*: The experiment has a mechanism that prevents two sounds from playing at the same time. If there are fewer than 10 ms since the previous sound stopped playing, the experiment will wait before playing the next sound. This prevents overlapping sounds when participants make a longe saccade and trigger multiple sounds.

```{r, echo= FALSE}
hist(s2$ISI-50, xlab= "ISI between two consecutive sounds (in ms)", main= "ISI timing", family= "serif", col="lightgreen", freq=FALSE, breaks=30, xlim= range(s2$ISI, na.rm=T))
abline(v = mean(s2$ISI-50, na.rm=T), col= "darkred", lwd=3)
curve(dnorm(x, mean=mean(s2$ISI-50, na.rm=T), sd=sd(s2$ISI-50, na.rm=T)), add=TRUE, col="darkblue", lwd=2) 

```

## Descriptive statistics:


### First two fixations after hearing sound:

```{r, echo= FALSE}
library(reshape)
library(ggplot2)

DesFix<- melt(sound_check, id=c('sub', 'item', 'cond', 'sound_type'), 
                           measure=c("N1", "N2"), na.rm=TRUE)
mFix<- cast(DesFix, sound_type ~ variable
                        ,function(x) c(M=signif(mean(x),3)
                              , SD= sd(x) ))
db<- data.frame(c(mFix$N1_M, mFix$N2_M), c(mFix$N1_SD, mFix$N2_SD),
                c(rep("First fixation",3),  rep("Second fixation",3)), 
                rep(c("Deviant", "Silence", "Standard"),2))

colnames(db)<- c("Mean", "SD", "Measure", "Sound")
db$SE<- db$SD/sqrt(length(unique(sound_check$sub)))
db$Sound<- factor(db$Sound, levels= c("Silence", "Standard", "Deviant"))

limits <- aes(ymax = db$Mean + db$SE, ymin=db$Mean - db$SE)

Dplot2<- ggplot(data= db, aes(x=Sound, y= Mean, color=Measure, fill= Measure, group= Measure, shape= Measure, linetype= Measure))+ 
  scale_fill_brewer(palette="Dark2")+ 
  scale_colour_brewer(palette="Dark2")+
  #scale_fill_manual(values=c(ColorBlind[6:7]))+
  #scale_colour_manual(values=c(ColorBlind[6:7]))+
  theme_bw() + theme(panel.grid.major = element_line(colour = "#E3E5E6", size=0.7), 
                     axis.line = element_line(colour = "black", size=1),
                     panel.border = element_rect(colour = "black", size=1, fill = NA))+
  geom_line(size=2)+
  geom_point(size=7)+ 
  xlab("\n Background sound")+ ylab("Mean fixation duration (in ms)")+ 
  theme(legend.position= "bottom", legend.title=element_text(size=20,
        face="bold", family="serif"),
        legend.text=element_text(size=20,family="serif"),legend.key.width=unit(2,"cm"),
        legend.key.height=unit(1,"cm"), strip.text=element_text(size=20, family="serif"),
        title=element_text(size=20, family="serif"),
        axis.title.x = element_text(size=20, face="bold", family="serif"),
        axis.title.y = element_text(size=20, face="bold", family="serif"), 
        axis.text=element_text(size=20, family="serif"), 
        panel.border = element_rect(linetype = "solid", colour = "black"), 
        legend.key = element_rect(colour = "#000000", size=1))+
  geom_ribbon(limits, alpha=0.1, colour=NA)


```

```{r, echo=FALSE, fig.height= 7, fig.width=9}
Dplot2
```


### Fixation duration measures (on the target word):

```{r, echo=FALSE}
if(office){
  load("C:/Users/mvasilev/Documents/DEVS/data/FD.Rda")
}else{
  load("C:/Users/Martin Vasilev/Documents/DEVS/data/FD.Rda")
}

#FD<- subset(FD, sub!=5)

library(reshape)
FD<- subset(FD,!is.na(sound))
FD$sound<- as.factor(FD$sound)
#tw<- subset(FD, is.element(word, c(3,5,7,9,11)))

DesFix<- melt(FD, id=c('sub', 'item', 'cond', 'sound'), 
              measure=c("FFD", "SFD", "GD", "TVT"), na.rm=TRUE)
mFix<- cast(DesFix, sound ~ variable
            , function(x) c(M=signif(mean(x),3)
                            , SD= sd(x) ))


library(ggplot2)

db<- data.frame(c(mFix$FFD_M, mFix$SFD_M, mFix$GD_M, mFix$TVT_M), c(mFix$FFD_SD, mFix$SFD_SD, mFix$GD_SD, mFix$TVT_SD),
                c(rep("FFD",3), rep("SFD",3),  rep("GD",3), rep("TVT",3)), 
                rep(c("Deviant", "Silence", "Standard"),4))

colnames(db)<- c("Mean", "SD", "Measure", "Sound")
db$SE<- db$SD/sqrt(length(unique(FD$sub)))

db$Sound<- factor(db$Sound, levels= c("Silence", "Standard", "Deviant"))

limits <- aes(ymax = db$Mean + db$SE, ymin=db$Mean - db$SE)

Dplot<- ggplot(data= db, aes(x=Sound, y= Mean, color=Measure, fill= Measure, group= Measure, shape= Measure, linetype= Measure))+ 
  scale_fill_brewer(palette="Dark2")+ 
  scale_colour_brewer(palette="Dark2")+
  #scale_fill_manual(values=c(ColorBlind[6:7]))+
  #scale_colour_manual(values=c(ColorBlind[6:7]))+
  theme_bw() + theme(panel.grid.major = element_line(colour = "#E3E5E6", size=0.7), 
                     axis.line = element_line(colour = "black", size=1),
                     panel.border = element_rect(colour = "black", size=1, fill = NA))+
  geom_line(size=2)+
  geom_point(size=7)+ 
  xlab("\n Background sound")+ ylab("Mean fixation duration (in ms)")+ 
  theme(legend.position= "bottom", legend.title=element_text(size=20,
        face="bold", family="serif"),
        legend.text=element_text(size=20,family="serif"),legend.key.width=unit(2,"cm"),
        legend.key.height=unit(1,"cm"), strip.text=element_text(size=20, family="serif"),
        title=element_text(size=20, family="serif"),
        axis.title.x = element_text(size=20, face="bold", family="serif"),
        axis.title.y = element_text(size=20, face="bold", family="serif"), 
        axis.text=element_text(size=20, family="serif"), 
        panel.border = element_rect(linetype = "solid", colour = "black"), 
        legend.key = element_rect(colour = "#000000", size=1))#+
  #geom_ribbon(limits, alpha=0.1, colour=NA)

```

```{r, echo=FALSE, fig.height= 7, fig.width=9}
Dplot

```
FFD: First fixation duration. SFD: Single fixation duration (i.e., the word was fixated only once during the trial). GD: gaze duration (the sum of all fixations before moving on to another word). TVT: Total viewing time (the sum of all fixations, including the ones made during a regression).

### Fixation durations on the NEXT word in the sentence after playing a sound:

```{r, echo= FALSE}
if(office){
  load("C:/Users/mvasilev/Documents/DEVS/data/N1.Rda")
}else{
  load("C:/Users/Martin Vasilev/Documents/DEVS/data/N1.Rda")
}

#N1<- subset(N1, sub!=5)

DesN1<- melt(N1, id=c('sub', 'item', 'cond', 'sound'), 
              measure=c("FFD", "SFD", "GD", "TVT"), na.rm=TRUE)
mN1<- cast(DesN1, sound ~ variable
            ,function(x) c(M=signif(mean(x),3)
                           , SD= sd(x) ))

db<- data.frame(c(mN1$FFD_M, mN1$SFD_M, mN1$GD_M, mN1$TVT_M), c(mN1$FFD_SD, mN1$SFD_SD, mN1$GD_SD, mN1$TVT_SD),
                c(rep("FFD",3), rep("SFD",3),  rep("GD",3), rep("TVT",3)), 
                rep(c("Deviant", "Silence", "Standard"),4))

colnames(db)<- c("Mean", "SD", "Measure", "Sound")
db$SE<- db$SD/sqrt(length(unique(N1$sub)))

db$Sound<- factor(db$Sound, levels= c("Silence", "Standard", "Deviant"))

limits <- aes(ymax = db$Mean + db$SE, ymin=db$Mean - db$SE)

Dplot<- ggplot(data= db, aes(x=Sound, y= Mean, color=Measure, fill= Measure, group= Measure, shape= Measure, linetype= Measure))+ 
  scale_fill_brewer(palette="Dark2")+ 
  scale_colour_brewer(palette="Dark2")+
  #scale_fill_manual(values=c(ColorBlind[6:7]))+
  #scale_colour_manual(values=c(ColorBlind[6:7]))+
  theme_bw() + theme(panel.grid.major = element_line(colour = "#E3E5E6", size=0.7), 
                     axis.line = element_line(colour = "black", size=1),
                     panel.border = element_rect(colour = "black", size=1, fill = NA))+
  geom_line(size=2)+
  geom_point(size=7)+ 
  xlab("\n Background sound")+ ylab("Mean fixation duration (in ms)")+ 
  theme(legend.position= "bottom", legend.title=element_text(size=20,
        face="bold", family="serif"),
        legend.text=element_text(size=20,family="serif"),legend.key.width=unit(2,"cm"),
        legend.key.height=unit(1,"cm"), strip.text=element_text(size=20, family="serif"),
        title=element_text(size=20, family="serif"),
        axis.title.x = element_text(size=20, face="bold", family="serif"),
        axis.title.y = element_text(size=20, face="bold", family="serif"), 
        axis.text=element_text(size=20, family="serif"), 
        panel.border = element_rect(linetype = "solid", colour = "black"), 
        legend.key = element_rect(colour = "#000000", size=1))#+
  #geom_ribbon(limits, alpha=0.1, colour=NA)
```

```{r, echo=FALSE, fig.height= 7, fig.width=9}
Dplot

```
FFD: First fixation duration. SFD: Single fixation duration (i.e., the word was fixated only once during the trial). GD: gaze duration (the sum of all fixations before moving on to another word). TVT: Total viewing time (the sum of all fixations, including the ones made during a regression).

### Number of fixations:

First-pass fixations are made during the initial reading of the word (they count towards GD). Second-pass fixations are those made during regressions to the word. When the two types of fixations are combined, they count towards TVT (i.e., the difference from GD is that we also add the second-pass fixations in TVT).

```{r, echo= FALSE}
if(office){
  load("C:/Users/mvasilev/Documents/DEVS/data/FixN.Rda")
  load("C:/Users/mvasilev/Documents/DEVS/data/FixN1.Rda")
}else{
  load("C:/Users/Martin Vasilev/Documents/DEVS/data/FixN.Rda")
  load("C:/Users/Martin Vasilev/Documents/DEVS/data/FixN1.Rda")
}

#FixN<- subset(FixN, sub!=5)
#FixN1<- subset(FixN1, sub!=5)

DesFixN<- melt(FixN, id=c('sub', 'item', 'cond', 'sound'), 
              measure=c("nfix1", "nfix2", "nfixAll"), na.rm=TRUE)
mFixN<- cast(DesFixN, sound ~ variable
            , function(x) c(M=signif(mean(x),3)
                            , SD= sd(x) ))

DesFixN1<- melt(FixN1, id=c('sub', 'item', 'cond', 'sound'), 
              measure=c("nfix1", "nfix2", "nfixAll"), na.rm=TRUE)
mFixN1<- cast(DesFixN1, sound ~ variable
            , function(x) c(M=signif(mean(x),3)
                            , SD= sd(x) ))
mFixN[,2:7]<- round(mFixN[,2:7],2)
mFixN1[,2:7]<- round(mFixN1[,2:7],2)

colnames(mFixN)<- c("Sound", "1st-pass (M)", "1st-pass (SD)", "2nd-pass (M)", "2nd-pass (SD)",
                    "All (M)", "All (SD)")

colnames(mFixN1)<- c("Sound", "1st-pass (M)", "1st-pass (SD)", "2nd-pass (M)", "2nd-pass (SD)",
                    "All (M)", "All (SD)")
mFixN$Sound<-  c("Silence", "Standard", "Deviant")
mFixN1$Sound<-  c("Silence", "Standard", "Deviant")
#levels(mFixN$sound)<- c("Silence", "Standard", "Deviant")
#levels(mFixN1$sound)<- c("Silence", "Standard", "Deviant")

library(knitr)
```


_On the target word:_

`r kable(mFixN)`

_On the next word in the sentence:_

`r kable(mFixN1)`